\section{Choosing Parameters}\label{sec:params}

\begin{table}
\begin{tabular}{c|c|c|c|c|c}
        &   $n$             &   $t_a$           &   $m$             &   $e$             &   $s$             \\
    \hline
    $A$ &   -               &   $\downarrow$    &   $\uparrow$      &   -               &   $\downarrow$    \\
    $H$ &   $\downarrow$    &   $\uparrow$      &   $\downarrow$    &   -               &   $\uparrow$      \\
    $P$ &   *               &   *               &   -               &   -               &   -               \\
    $S$ &   -               &   $\uparrow$      &   -               &   -               &   -               \\
    $F$ &   -               &   -               &   $\downarrow$    &   -               &   $\downarrow$    \\
    $G$ &   -               &   -               &   $\downarrow$    &   $\downarrow$    &   -
\end{tabular}
\end{table}

\begin{table}
\begin{tabular}{c|c|c|c|c|c|c}
            &   $A$             &   $H$             &   $P$ &   $S$ &   $F$ &   $G$ \\
    \hline
    $n_*$   &   -               &   $\downarrow$    &   -   &   -   &   -   &   -   \\
    $d_*$   &   $\downarrow$    &   -               &   -   &   -   &   -   &   -
\end{tabular}
\end{table}

\begin{table}
\begin{tabular}{c|c|c|c|c|c}
            &   $n$             &   $t_a$           &   $m$             &   $e$             &   $s$             \\
    \hline
    $n_*$   &   $\uparrow$      &   $\downarrow$    &   $\uparrow$      &   -               &   $\downarrow$    \\
    $d_*$   &   -               &   $\uparrow$      &   $\downarrow$    &   -               &   $\uparrow$      \\
\end{tabular}
\end{table}

Parameter selection is important for smooth and interesting gameplay.
One should consider a number of options when deciding on parameter values.

\subsection{The $P$ Parameter}

The $P$ parameter is the most nebulous of the parameters,
as it doesn't correspond to any real statistic.
The current recommendation is to assign it a value of 2.

\subsection{The $H$ Parameter}

The $H$ parameter represents the health of the units,
and directly scales their number of dice, which act as hitpoints.

\subsubsection{Minimum Dice}

In order to ensure that each unit has a reasonable number of dice,
it may be desirable to choose a value $n_*$
such that every unit has at least $n_*$ dice.

The unit with the least dice is that with the lowest health,
calculated according to the health factor:
\[
    h u r^P
\]
From the statistics of the least-health unit, the formula for $H$
can easily be derived using the normal $n$ formula.
\[
    n_* =
        \frac
            {h u r^P}
            {10^P H}
\]
can be rearranged to
\[
    H =
        \frac
            {h u r^P}
            {10^P n_*}
\]

\todo[inline]{example}

\subsection{The $A$ Parameter}

The $A$ parameter defines a ``standard'' amount of damage:
the amount of damage a unit must do to reduce an opponent's dice by one.
A low $A$ value can speed up combat,
and may reduce the relative strength of the strongest units;
as hit chance cannot exceed 100\%.
A high $A$ value can slow down combat,
and may leave the weakest units with no chance to hit.
(Units that cannot hit are referred to as \emph{fodder}.)

\subsubsection{Median Damage}

Having already determined the $H$ parameter and
therefore the number of dice of each unit,
may be useful to select $A$ by the average damage of an attack.\footnote{
    Note that the maximum and minimum damage of an attack
    are already determined by the number of dice of each unit,
    as the most damage on unit can do is equal to its health.
}

We begin by identifying the median attack,
according to the attack-defend ratio calculated by the following formula.
\[
    \frac
        {n d_a (10 + b_a)^P}
        {h r^P}
\]
With the median attack in hand,
we can use the unit's statistics to derive an appropriate $A$ value
such that this attack is most likely to deal our target average damage, $d_*$.
Recall the previously introduced hit chance formula:
\[
    p = \frac
        {d_a (10 + b_a)^P H}
        {h r^P A}
\]
We want to choose a value $A$ such that $d_* = n p$:
\[
    d_* = \frac
        {n d_a (10 + b_a)^P, H}
        {h r^P A}
\]
We rearrange to get the formula:
\[
    A = \frac
        {n d_a (10 + b_a)^P, H}
        {h r^P d_*}
\]

\todo[inline]{pick A by min target $t$}
\subsection{The $S$ Parameter}
The $S$ parameter is generally set to 20 in this document,
as it balances granularity and simplicity,
and is familiar to d20 system users.
It is possible, however, that greater granularity is needed.
This might arise in cases where,
after calculating an appropriate $A$ value,
one or more battle participants needs only to roll 1 to hit,
(i.e., will always hit),
or otherwise if participants whose strength ought to be different
identical statistics.

These issues may be addressed by increasing the $S$ parameter to 100,
for example.
However, they may instead point to an issue that some unit
is so weak as to be in ineffectual and might be disregarded for the battle.
