\section{Derivation}

\subsubsection{Accuracy}

D\&D's concept of accuracy is a source of complications in the formulas,
and must be addressed specifically.

We begin by observing that in standard D\&D rules,
an increase by a factor of $x$ in an attack's accuracy
can be roughly equivalently represented by a decrease in a defender's
health by a factor of $x$.
For example, if an attack normally has a 25\% chance of hitting a defender with 2 hitpoints,
we might reasonably expect the defender to last 8 rounds.\footnote{
    This approximation follows intuition,
    but is not statistically correct with respect to expected values.
    It is considered sufficient for this draft
    but may be revisited in a future version.
}
If we raise the accuracy to 50\%,
the defender would need around 4 hitpoints to last 8 rounds.

Accuracy of attacks in standard D\&D is not derived from one statistic,
but from two: the attacker's attack bonus and the defender's AC.
Modeling this directly would present a problem for the transformation,
as it would necessitate an additional statistic and comparison during gameplay.
Instead, we try to bundle the effect of AC and health into $n$,
and the effect of attack bonus into $t_a$.

It is not possible to replicate the effect of accuracy fully, however.
To demonstrate, we turn to an example of a commoner (+2 attack bonus)
attacking a tarrasque (25 AC).
Barring critical hits, the commoner will never be able to hit the tarrasque,
as the maximum roll of 22 is less than the tarrasque's AC.
The result is that the $n$ value for a tarrasque unit must be infinite,
or that the $t_a$ value for a commoner unit must be at least 20.

Clearly, neither of these options is practical.
Instead of modeling the asymptotic effect of AC and attack bonus directly,
we use power function to approximate it.
AC is converted into a defense factor:
\[
    \left(\frac{r}{10}\right)^P
\]
AC is first divided by 10,
as 10 is the standard AC without bonuses or penalties,
then raised to a power $P$ which represents the multiplicative power of AC.

The factor for attack bonus is derived similarly,
but 10 added, establishing 0 as the baseline attack bonus:
\[
    \left(\frac{10 + b_a}{10}\right)^P
\]

\subsubsection{Deriving the Formulas}

The $n$ formula is relatively simple:
It is the total health of the unit,
multiplied by a factor representing the value of armor,
divided by the $H$ parameter:
\[
    \frac
        {h u (\frac{r}{10})^P}
        {H}
\]
We round and simplify to get our value for $n$.

The $t$ formula is more complex,
as it must take into account both attacking strength and defensive strength.
Because $n$ dice (proportional to defense) will be rolled,
in order to avoid a compounding impact of defense,
the defense must be divided out of the equation.

The development of the formula is as follows:
The attack value \emph{per creature} is proportional to
the average damage multiplied by the attack bonus factor.
\[
    d_a \left(\frac{10 + b_{a}}{10}\right)^{P}
\]
The total attack value of the unit is then
\[
    d_a u \left(\frac{10 + b_{a}}{10}\right)^{P}
\]
We divide this by the number of dice to get attack value per die:
\[
    \frac
        {d_a u \left(\frac{10 + b_{a}}{10}\right)^{P}}
        {n}
\]
But to reduce rounding error we instead use the unrounded formula for $n$:
\[
    \frac
        {d_a u \left(\frac{10 + b_{a}}{10}\right)^{P}}
        {\frac{h u r^P}{10^P H}}
\]
which simplifies to
\[
    \frac
        {d_a (10 + b_{a})^P H}
        {h r^P}
\]
This value is proportional to our hit chance.
We scale it to a probability by dividing by $A$:
\[
    \frac
        {d_a (10 + b_{a})^P H}
        {h r^P A}
\]
We now have our true hit chance.
To turn this into a target roll,
we multiply by $S$, round, subtract from $S$ and add 1.
\[
    S -
    \left\lfloor
        \frac
            {d_a (10 + b_{a})^P H S}
            {h r^P A}
    \right\rceil
    + 1
\]


\section{Choosing Parameters}\label{sec:params}

\begin{table}
\begin{tabular}{c|c|c|c|c|c}
        &   $n$             &   $t_a$           &   $m$             &   $e$             &   $s$             \\
    \hline
    $A$ &   -               &   $\downarrow$    &   $\uparrow$      &   -               &   $\downarrow$    \\
    $H$ &   $\downarrow$    &   $\uparrow$      &   $\downarrow$    &   -               &   $\uparrow$      \\
    $P$ &   *               &   *               &   -               &   -               &   -               \\
    $S$ &   -               &   $\uparrow$      &   -               &   -               &   -               \\
    $F$ &   -               &   -               &   $\downarrow$    &   -               &   $\downarrow$    \\
    $G$ &   -               &   -               &   $\downarrow$    &   $\downarrow$    &   -
\end{tabular}
\end{table}

\begin{table}
\begin{tabular}{c|c|c|c|c|c|c}
            &   $A$             &   $H$             &   $P$ &   $S$ &   $F$ &   $G$ \\
    \hline
    $n_*$   &   -               &   $\downarrow$    &   -   &   -   &   -   &   -   \\
    $d_*$   &   $\downarrow$    &   -               &   -   &   -   &   -   &   -
\end{tabular}
\end{table}

\begin{table}
\begin{tabular}{c|c|c|c|c|c}
            &   $n$             &   $t_a$           &   $m$             &   $e$             &   $s$             \\
    \hline
    $n_*$   &   $\uparrow$      &   $\downarrow$    &   $\uparrow$      &   -               &   $\downarrow$    \\
    $d_*$   &   -               &   $\uparrow$      &   $\downarrow$    &   -               &   $\uparrow$      \\
\end{tabular}
\end{table}

Parameter selection is important for smooth and interesting gameplay.
One should consider a number of options when deciding on parameter values.

\subsection{The $P$ Parameter}

The $P$ parameter is the most nebulous of the parameters,
as it doesn't correspond to any real statistic.
The current recommendation is to assign it a value of 2.

\subsection{The $H$ Parameter}

The $H$ parameter represents the health of the units,
and directly scales their number of dice, which act as hitpoints.

\subsubsection{Minimum Dice}

In order to ensure that each unit has a reasonable number of dice,
it may be desirable to choose a value $n_*$
such that every unit has at least $n_*$ dice.

The unit with the least dice is that with the lowest health,
calculated according to the health factor:
\[
    h u r^P
\]
From the statistics of the least-health unit, the formula for $H$
can easily be derived using the normal $n$ formula.
\[
    n_* =
        \frac
            {h u r^P}
            {10^P H}
\]
can be rearranged to
\[
    H =
        \frac
            {h u r^P}
            {10^P n_*}
\]

\todo[inline]{example}

\subsection{The $A$ Parameter}

The $A$ parameter defines a ``standard'' amount of damage:
the amount of damage a unit must do to reduce an opponent's dice by one.
A low $A$ value can speed up combat,
and may reduce the relative strength of the strongest units;
as hit chance cannot exceed 100\%.
A high $A$ value can slow down combat,
and may leave the weakest units with no chance to hit.
(Units that cannot hit are referred to as \emph{fodder}.)

\subsubsection{Median Damage}

Having already determined the $H$ parameter and
therefore the number of dice of each unit,
may be useful to select $A$ by the average damage of an attack.\footnote{
    Note that the maximum and minimum damage of an attack
    are already determined by the number of dice of each unit,
    as the most damage on unit can do is equal to its health.
}

We begin by identifying the median attack,
according to the attack-defend ratio calculated by the following formula.
\[
    \frac
        {n d_a (10 + b_a)^P}
        {h r^P}
\]
With the median attack in hand,
we can use the unit's statistics to derive an appropriate $A$ value
such that this attack is most likely to deal our target average damage, $d_*$.
Recall the previously introduced hit chance formula:
\[
    p = \frac
        {d_a (10 + b_a)^P H}
        {h r^P A}
\]
We want to choose a value $A$ such that $d_* = n p$:
\[
    d_* = \frac
        {n d_a (10 + b_a)^P, H}
        {h r^P A}
\]
We rearrange to get the formula:
\[
    A = \frac
        {n d_a (10 + b_a)^P, H}
        {h r^P d_*}
\]

\todo[inline]{pick A by min target $t$}

\subsection{The $S$ Parameter}
The $S$ parameter is generally set to 20 in this document,
as it balances granularity and simplicity,
and is familiar to d20 system users.
It is possible, however, that greater granularity is needed.
This might arise in cases where,
after calculating an appropriate $A$ value,
one or more battle participants needs only to roll 1 to hit,
(i.e., will always hit),
or otherwise if participants whose strength ought to be different
identical statistics.

These issues may be addressed by increasing the $S$ parameter to 100,
for example.
However, they may instead point to an issue that some unit
is so weak as to be in ineffectual and might be disregarded for the battle.

\subsection{The $F$ Parameter}

\subsubsection{Scaling Time}

\[
    F = \sqrt{u_\rightarrow}
\]
We can calculate the time of round via the $F$ parameter,
which represents the fraction of a unit that is actually attacking on a given turn.
\[
    s = \frac
        {3 A}
        {H F}
\]
\todo[inline]{can calculate via assuming only periphery will attack, etc.}
\todo[inline]{explain why div by 2}
\todo[inline]{explain why mul by 3 (6 / 2)}
\todo[inline]{explain s}

\color{red}
\[
    s = \frac{12 A}{H F}
\]
\color{black}

\subsection{The $G$ Parameter}

\subsubsection{Scaling Space}

Each unit takes up space according to its members:
a unit of 20 medium creatures takes a total of 500 square feet (25 square feet per creature).
When playing on a grid,
some rounding must be done to ensure that each unit conforms to the grid;
i.e., that each unit's tile size is a square number.

We scale the size of units relative to the grid
according to parameter $G$,
which indicates the length of the edge of a grid tile.
A unit's edge length is calculated by the following:
\[
    e = \frac
        {\sqrt{c^2 u}}
        {G}
\]

\subsubsection{Scaling Speed}
The marching speed $m$ of a unit is determined
by the scaling of time and space
as applied to the creature's speed.
We begin with the creature's speed and divide by 6
to find the distance traveled per second.
\[
    \frac
        {p}
        {6}
\]
We multiply by the number of seconds per round:
\[
    \frac
        {p}
        {6}
    \left(
        \frac
            {A}
            {2 H F}
    \right)
\]
And finally we divide by the number of feet per tile,
simplifying to
\[
    m = \frac
        {p A}
        {12 H F G}
\]

\subsubsection{Minimum Unit Size}

One may select $G$ in such a way that the physically smallest unit
occupies exactly one square.

First, the smallest unit must be found,
which is the unit for which the unscaled size value is smallest:
\[
    \sqrt{c^2 u}
\]
We then select $G$ such that $e = 1$;
i.e., we solve:
\[
    G = \frac
            {\sqrt{c^2 u}}
            {1}
\]
which is easily simplified to
\[
    G = \sqrt{c^2 u}
\]

\subsubsection{Minimum Unit Speed}

One may selection $G$ in such a way that the slowest unit moves one square per turn.

First, the slowest unit must be found,
which is the unit whose creature speed value $p$ is least.

We rearrange the formula for $m$ in order to find the $G$ value:
\[
    G = \frac
            {p A}
            {12 H F m}
\]
Setting $m$ to 1 tile, we have:
\[
    G = \frac
            {p A}
            {12 H F}
\]

\subsubsection{Attack Range}

With the size of the grid established, the attack range for the units may be established.
This is done simply by taking the long range for an attack,
dividing by the grid size,
and rounding up.

\[
    x_a = \left\lceil\frac
        {l_a}
        {G}
        \right\rceil
\]

\todo[inline]{Mix the two: take $G = min(G_1, G_2)$}